Combining information from photometric and spectroscopic data:\\ Don't waste your time in spectrophotometric calibration!

### 2014-04-26 tex/ms.tex
comments
tweaks

% This file is part of the speccal project.
% Copyright 2014 the authors.

% ## style notes:
% - all equations in `eqnarray` environments
% - parameters are `inferred`; `measurement` is reserved for data

% \documentclass[iop,numberedappendix]{emulateapj}
% \usepackage{apjfonts}
\documentclass[12pt, letterpaper, preprint]{aastex}
\bibliographystyle{apj}

\input{vc}
\newcommand{\transpose}[1]{{#1}^{\!\mathsf T}}
\newcommand{\given}{\,|\,}

\newcommand{\lick}{1}
\newcommand{\ccpp}{2}
\newcommand{\cds}{3}
\newcommand{\mpia}{4}

\begin{document}

\title{Combining information from photometric and spectroscopic data:\\
  Don't waste your time in spectrophotometric calibration!}
\author{
  BDJ\altaffilmark{\lick},
  DRW\altaffilmark{\lick},
  DFM\altaffilmark{\ccpp},
  DWH\altaffilmark{\ccpp, \cds, \mpia},
  CC\altaffilmark{\lick},
  others}
\altaffiltext{\lick}{University of California Lick Observatory}
\altaffiltext{\ccpp}{Center for Cosmology and Particle Physics, Department of Physics, New York University}
\altaffiltext{\cds}{Center for Data Science, New York University}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie}

\begin{abstract}
In a typical spectral-fitting project,
  there tend to be a small number of bands of well-calibrated photometry
  and thousands of poorly calibrated spectroscopic pixels.
Good spectroscopic calibration is both expensive and difficult,
  but naive weighted least-square fitting will weight
  the thousands of spectroscopic pixels far higher overall
  than the few photometric bands.
Here we present a flexible and general spectroscopic calibration model,
  which can be fit simultaneously along with whatever spectral parameters.
The model involves a polynommial calibration vector to capture large-scale calibration issues
  and a Gaussian Process to capture small-scale wiggles.
We show that, in this framework, the quality of some kinds of spectral parameter fits
  is not a strong function of the quality of the spectroscopic calibration;
  that is, for many scientific goals there is no scientific reason
  to obtain good---or really any---spectrophotometric calibration.
In particular, for stellar population fitting, high signal-to-noise data,
  and good photometry,
  we show that raw counts are just as good as calibrated counts for making inferences about the
  parameters of interest.
This obviation of calibration overheads will pay all yall \emph{mad cheddar},
  at least if all yall employ full cost accounting.
\end{abstract}

@@

The Right Thing To Do (RTTD) is to express the \emph{reason} that the
RPI believes the spectroscopic data to be less reliable \emph{within}
the likelihood function.
That is, the RPI must add nuisance parameters to parameterize the
unreliability of the spectroscopy, and infer those along with the
parameters of interest.
If this parameterization is good, the information in the spectroscopy
data will not flow entirely to the parameters of interest; instead,
some will flow to the nuisance parameters.
This will reduce the importance of the spectroscopy relative to the
photometry and the RPI will be well pleased with the dominance of the
photometric data in the final inferences of the parameters of
interest.
In this what follows, we provide an example of this kind of
parameterization and inference.

@@

In detail, we are going to choose a $M$th-order polynomial form for
the calibration vector $f(\alpha)$ and a bell-shape (Gaussian) for the
kernel function $K_\alpha(\cdot)$:
\begin{eqnarray}
f(\alpha) &=& \sum_{m=1}^M c_m\,x^m
\\
x &\equiv& \frac{\lambda}{\lambda_0} - 1
\\
K_\alpha(x) &=& a^2\,\exp(-\frac{x^2}{2\,\ell^2})
\\
\transpose{\alpha} &\equiv& \left[ \{c_m\}_{m=0}^M, a^2, \ell^2 \right]
\quad ,
\end{eqnarray}
where $\lambda$ is a column vector made up of the wavelengths $\lambda_m$
and $x$ is a scaled and shifted version of the same,
scaled by a fiducial wavelength $\lambda_0$.
Nuisance parameters $c_m$ set the shape of the large-scale calibration
issues, while $a^2$ and $\ell^2$ set the variance and correlation
length of the remaining covariant calibration noise.

>>>>

% This file is part of the `speccal` project.
% Copyright 2014 the authors.
% All rights reserved.

% ## style notes:
% - all equations in `eqnarray` environments
% - parameters are `inferred`; `measurement` is reserved for data
% - use emulateapj or aastex, as you like, but watch for equation-wrapping in the former

% ---- Hogg (aastex) mode ----
% \documentclass[12pt, letterpaper, preprint]{aastex}
% ----

% ---- Charlie (emulateapj) mode ----
\documentclass[iop,numberedappendix]{emulateapj}
\usepackage{apjfonts}
% ----

\bibliographystyle{apj}
\input{vc}

% math shih
\newcommand{\transpose}[1]{{#1}^{\!\mathsf T}}
\newcommand{\given}{\,|\,}

% affiliation marks; edit with care
\newcommand{\lick}{1}
\newcommand{\ccpp}{2}
\newcommand{\cds}{3}
\newcommand{\mpia}{4}

\begin{document}\sloppy\sloppypar
\title{Combining information from photometric and spectroscopic data:\\
  Don't waste your time in spectrophotometric calibration!}
\shorttitle{don't waste your time in calibration!}
\author{
  BDJ\altaffilmark{\lick},
  DRW\altaffilmark{\lick},
  DFM\altaffilmark{\ccpp},
  DWH\altaffilmark{\ccpp, \cds, \mpia},
  CC\altaffilmark{\lick},
  others}
\shortauthors{bdj et al}
\altaffiltext{\lick}{University of California Lick Observatory}
\altaffiltext{\ccpp}{Center for Cosmology and Particle Physics, Department of Physics, New York University}
\altaffiltext{\cds}{Center for Data Science, New York University}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie}

\begin{abstract}
In a typical spectral-fitting project,
  there tend to be a small number of bands of well-calibrated photometry
  and thousands of poorly calibrated spectroscopic pixels.
Good spectroscopic calibration is both expensive and difficult,
  but naive weighted least-square fitting will weight
  the thousands of spectroscopic pixels far higher overall
  than the few photometric bands.
Here we present a flexible and general spectroscopic calibration model,
  which can be fit simultaneously along with whatever spectral parameters.
The model involves a polynommial calibration vector to capture large-scale calibration issues
  and a Gaussian Process to capture small-scale wiggles.
We show that, in this framework, the quality of some kinds of spectral parameter fits
  is not a strong function of the quality of the spectroscopic calibration;
  that is, for many scientific goals there is no scientific reason
  to obtain good---or really any---spectrophotometric calibration.
In particular, for stellar population fitting, high signal-to-noise data,
  and good photometry,
  we show that raw counts are just as good as calibrated counts for making inferences about the
  parameters of interest.
This obviation of calibration overheads will pay \emph{all yall mad cheddar},
  at least if yall employ full cost accounting.
\end{abstract}

@@

The Right Thing To Do is to express the \emph{reason} that the RPI
believes the spectroscopic data to be less reliable \emph{within} the
likelihood function.
That is, the RPI must add nuisance parameters to parameterize the
unreliability of the spectroscopy, and infer those along with the
parameters of interest.
If this parameterization is good, the information in the spectroscopy
data will not flow entirely to the parameters of interest; instead,
some will flow to the nuisance parameters.
This will reduce the importance of the spectroscopy relative to the
photometry and the RPI will be well pleased with the dominance of the
photometric data in the final inferences of the parameters of
interest.
In this what follows, we provide an example of this kind of
parameterization and inference.

@@

In detail, we are going to choose a $M$th-order polynomial form for
the calibration vector $f(\alpha)$ and a bell shape (Gaussian) for the
kernel function $K_\alpha(\cdot)$:
\begin{eqnarray}
f(\alpha) &=& \sum_{m=1}^M c_m\,x^m
\\
x &\equiv& \frac{\lambda}{\lambda_0} - 1
\\
K_\alpha(x) &=& a^2\,\exp(-\frac{x^2}{2\,\ell^2})
\\
\transpose{\alpha} &\equiv& \left[ \{c_m\}_{m=0}^M, a^2, \ell^2 \right]
\quad ,
\end{eqnarray}
where $\lambda$ is a column vector made up of the wavelengths $\lambda_m$
and $x$ is a scaled and shifted version of the same,
scaled by a fiducial wavelength $\lambda_0$.
Nuisance parameters $c_m$ set the shape of the large-scale calibration
issues, while $a^2$ and $\ell^2$ set the variance and correlation
length of the remaining covariant calibration noise.
